Neural Power Unit \\-- A Bayesian Approach To Neural Arithmetic

### 2020-05-18 papers/main.tex
fuzz fuzz

% figures
\usepackage{tikz}
\usetikzlibrary{arrows,backgrounds,positioning,intersections}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{polar}
\usepgfplotslibrary{smithchart}
\usepgfplotslibrary{statistics}
\usepgfplotslibrary{dateplot}
\graphicspath{{./../plots/}}
\newcommand{\real}{\text{real}}

@@

@@

\begin{figure}
  \centering
  %\includegraphics[width=0.8\linewidth]{compare-npu-nmu-nalu-xy-xdivy.tikz}
  \input{../plots/compare-npu-nmu-nalu-xy-xdivy.tex}
  \caption{Comparison of different models learning the function
  $(x,y)\rightarrow(xy,\frac{x}{y})$ with $(x,y)\in\mathcal U^2(0.1,2)$.  Each
  plot shows the mean-squared error on a logarithmic color scale ($\hat{\bm
  t}=\text{model}(\bm x)$).  The Dense network learns both multiplication and
  division, but fails to extrapolate. NALU learns both tasks perfectly, but
  completely fails if one of the inputs is negative.  The NMU successfully
  learns multiplication, but not division. Only the NPU learns both tasks and
  can extrapolate with an error that is less than $0.01$ in most places.}%
  \label{fig:compare-npu-nmu-nalu-xy-xdivy}
\end{figure}

@@

@@

The combination of NAU and NMU can thus learn $\{+,-,\times\}$, but no division.

\todo{put this in acknowledgements
all the results in this paper were create with the help of the following Julia
packages: \cite{rackauckas_differentialequationsjl_2017} + (Flux.jl)}
\todo{add sqrt and positive range to table~\ref{tab:compare_npu_npu_nalu}?}
\begin{table}
  \centering
  \caption{Comparison of different $2\times2$ layers on the task
  $(x,y)\rightarrow(xy,\frac{x}{y})$.  $(x,y) \in \mathcal U^2(-2,2)$.}
  \label{tab:compare_npu_npu_nalu}
  \input{table-x*y-xdivy-pos-neg.tex}
\end{table}

@@

@@

Neural Arithmetic Units aim to overcome this problem with units that are able
to represent simple arithmetic operations exactly. By composition of these
units it becomes possible to learn e.g. exponentials, logarithms, and power
functions. This promises to improve on the extrapolation capabilities of NNs
even on tasks that are beyond artificial arithmetic tasks such as learning
multiplication, or a certain power function \todo{cite some examples}.
\todo{mention ODEs as use case}
\todo{mention explainability}

>>>>

% figures
\graphicspath{{./../plots/}}
\usepackage{tikz}
\usetikzlibrary{arrows,backgrounds,positioning,intersections}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{polar}
\usepgfplotslibrary{smithchart}
\usepgfplotslibrary{statistics}
\usepgfplotslibrary{dateplot}

\newcommand{\real}{\text{real}}

@@

Including Neural Arithmetic Units in common neural networks, apart from the
potential increase in extrapolation capabilities, has additional advantages.
They promise to reduce the amount of parameters that are needed for a given
task, can drastically improve the explainability of our models (more on the
explainability of the NPU in Sec.~\ref{sec:neural_power_unit}).  Additionally,
the field of Scientific Machine Learning, which, in part, is concerned with
learning differential equations~\citep{rackauckas_universal_2020} could benefit
from the inductive bias neural arithmetic.

@@

@@

\todo{should this be here already?}
With the simple exemplary task of learning a function $f:\mathbb R^2 \rightarrow \mathbb R^2$
\begin{equation}
  \label{eq:xx_xdivy}
  f(\bm x) = f(x,y) = [xy,x/y] = \bm t
\end{equation}
we can demonstrate the strengths and weaknesses of the NALU.  The NALU only
learns successfully if it is trained on positive numbers, so we train with
samples from $\bm x \in \mathcal U^2(0.1,2)$ and plot the error $\epsilon =
|\bm t - \hat{\bm t}|^2$ in Fig.~\ref{fig:compare_npu_nmu_nalu_xy_xdivy}.

@@

The combination of NAU and NMU can thus learn $\{+,-,\times\}$, but no division.
This is shown in Fig.~\ref{fig:compare_npu_nmu_nalu_xy_xdivy} as well. The NMU
can perfectly represent the multiplication of the two inputs, but fails at
learning division.

\todo{put this in acknowledgements
all the results in this paper were create with the help of the following Julia
packages: \cite{rackauckas_differentialequationsjl_2017} + (Flux.jl)}

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{compare-npu-nmu-nalu-xy-xdivy.png}
  \caption{Comparison of different models learning the function
  $(x,y)\rightarrow(xy,x/y)$ with $(x,y)\in\mathcal U^2(0.1,2)$.  Each
  plot shows the mean-squared error on a logarithmic color scale ($\hat{\bm
  t}=\text{model}(\bm x)$).  The Dense network learns both multiplication and
  division, but fails to extrapolate. NALU learns both tasks perfectly, but
  completely fails if one of the inputs is negative.  The NMU successfully
  learns multiplication, but not division. Only the NPU learns both tasks and
  can extrapolate with an error that is less than $0.01$ in most places.}%
  \label{fig:compare_npu_nmu_nalu_xy_xdivy}
\end{figure}
\todo{add in depth experiment desc. of fig.~\ref{fig:compare_npu_nmu_nalu_xy_xdivy} in appendix}
\todo{add sqrt to table~\ref{tab:compare_npu_npu_nalu}?}
\begin{table}
  \centering
  \caption{Comparison of different $2\times2$ layers on the task
  $(x,y)\rightarrow(xy,\frac{x}{y})$.  $(x,y) \in \mathcal U^2(-2,2)$.}
  \label{tab:compare_npu_npu_nalu}
  \input{table-x*y-xdivy-pos.tex}
\end{table}

@@

Eq.~\ref{eq:nalu_mult} enables the NALU to learn multiplication. If the weight
matrix $\bm W$ would not be restricted to values in the range $(-1,1)$ it could
in principle learn arbitrary power functions, because
\begin{align}
  z = \exp(w\log x) = x^w \text{ for } x>0.
\end{align}
However, this would complicate the convergence issues of the NALU even more.
We therefore suggest to leverage the power of the complex plane, by using the
complex logarithm and by promoting $w$ to a complex number.  This will enable
us not only to lift the positivity constraint on $x$, but also provides and
additional dimension that helps to avoid regions with an uninformative gradient
signal\todo{rephrase this. do we have an educating example where it is obvious
that complex plane helps? the stuff in fig.1 does not work as well with
non-complex NPU, but I have not idea how to plot that}.
Introducing a complex weight matrix somewhere in a larger network would mean that
the gradients in other layers would potentially become complex as well, which
would result in converting the whole network to complex numbers.
This seems like a high price to pay, but fortunately there is a way around this.
We can output only the real part of the result $z$, but still maintain a complex $w$.

\newcommand{\wre}{w_{\text{re}}}
\newcommand{\wim}{w_{\text{im}}}
\begin{align}
  \real(z) &= \real(\exp(w\log x)) \\
    &= \real(\exp((\wre + i\wim) (\log r + ik\pi))) \\
    &= \exp(\wre\log r - \wim k\pi) \real(\exp(i\wim\log r + ik\pi\wre)) \\
    &= \exp(\wre\log r - \wim k\pi) \cos(\wim\log r + k\pi\wre)
\end{align}
\todo{explicitly compute gradients?}

@@

Neural Arithmetic Units aim to overcome this problem with units that are able
to represent simple arithmetic operations exactly. By composition of these
units it becomes possible to learn e.g. exponentials, logarithms, and power
functions. This promises to improve on the extrapolation capabilities of NNs
even on tasks that are beyond artificial arithmetic tasks such as learning
multiplication, or a certain power function \todo{cite some examples}.
